---
title: "Boris Leroy's CV (short version)"
output:
  pdf_document: default
  md_document:
    variant: gfm
    preserve_yaml: true
layout: cv
---



# Boris Leroy
Maitre de Conférences, Muséum National d'Histoire Naturelle

Biology of Aquatic Organisms and Ecosystems<br/>
43 rue Cuvier, CP 26<br/>
75005 Paris, France<br/>

leroy.boris@@gmail.com



[borisleroy.com](https://borisleroy.com) - 
[lab page](https://borea.mnhn.fr/fr/users/boris-leroy) <br/>  
[GitHub: Farewe](https://github.com/farewe) - 
[ORCID: 0000-0002-7686-4302](https://orcid.org/0000-0002-7686-4302) - 
[RG: Boris_Leroy](https://www.researchgate.net/profile/Boris_Leroy) - 
[\@BLeroyEcology](https://twitter.com/BLeroyEcology)

```{r echo=FALSE, results="hide", message=FALSE, warning=FALSE}
library(scholar)
id <- "7HzlVT4AAAAJ"

gs <- get_profile(id)

```

*My research focuses on global biogeographical patterns of aquatic biodiversity (marine and freshwater) and their underlying direct and indirect drivers. I investigate with various macroecological and biogeographical methods how global changes alter the natural distribution of biodiversity. My main research topics are: biogeography, macroecology, invasion biology; with a special emphasis on the development of methods and facilitating access to these methods with open-source software. *

## Education
`2012`
**Ph.D., Life Sciences**, *Muséum national d'histoire naturelle/Université de Rennes 1*, Rennes, France

## Experience
`2014-present`
**Maître de conférences (lecturer)**, *Lab. Biology of Aquatic Organisms & Ecosystems, Muséum national d'Histoire Naturelle*, Paris, France.

`2014`
**Post-doctoral fellow**, *UMR 8079 Ecology Systematics & Evolution, University of Paris South XI*, Orsay, France.

`2013`
**Post-doctoral fellow**, *Lab. Biodiversity & Territory Management, University of Rennes 1*, Rennes, France.

## Publications

*[Full Publication list  available by clicking here](https://farewe.github.io/markdown-cv/publications.html).*

### Journal articles
56 published peer-reviewed articles. h-index: `r gs$h_index`; i10-index: `r gs$i10_index`

*Selected publications:*

**Leroy B**. **Leroy B**. 2023. Choosing presence-only species distribution models. *Journal of Biogeography* 50:247-250 [10.1111/jbi.14505](https://doi.org/10.1111/jbi.14505)  ---  [HAL](https://hal.archives-ouvertes.fr/hal-03813698)

Bernery C, Bellard C, Courchamp F, Brosse S, Gozlan R, Jarić I, Teletchea F, **Leroy B**.  2022. Freshwater fish invasions: A comprehensive review. *Annual Review of Ecology, Evolution and Systematics*  [10.1146/annurev-ecolsys-032522-015551](https://doi.org/10.1146/annurev-ecolsys-032522-015551) ---  [HAL](https://hal.archives-ouvertes.fr/hal-03781186)

**Leroy B**, Kramer AV, Vaissière AC, Kourantidou M, Courchamp F, Diagne C. 2022. Analysing global economic costs of invasive alien species with the invacost R package. *Methods in Ecology & Evolution* 13:1930-1937 [10.1111/2041-210X.13929](https://doi.org/10.1111/2041-210X.13929) ---  [HAL](https://hal.archives-ouvertes.fr/hal-03427555)

Schickele A, Goberville E, **Leroy B**, Beaugrand G, Hattab T, Francour P & Raybaud V. 2021. European small pelagic fish distribution under global change scenarios. *Fish and Fisheries* 22:212-225 [10.1111/faf.12515](https://doi.org/10.1111/faf.12515) ---  [HAL Link](https://HAL Link.archives-ouvertes.fr/HAL Link-03326581)

Diagne C, **Leroy B**, Vaissière AC, Gozlan RE, Roiz D, Jarić I, Salles JM, Bradshaw CJA & Courchamp F. 2021. High and rising economic costs of biological invasions worldwide *Nature* 592, pages 571–576 [10.1038/s41586-021-03405-6](https://doi.org/10.1038/s41586-021-03405-6)

**Leroy B**, Dias MS, Giraud E, Hugueny B, Jézéquel C,  Leprieur F, Oberdorff T, Tedesco PA. 2019. Global biogeographical regions of freshwater fish species. *Journal of Biogeography*, 46:2407-2419 [10.1111/jbi.13674](https://doi.org/10.1111/jbi.13674)

**Leroy B**, Gallon R, Feunteun E, Robuchon M & Ysnel F. 2017. Cross-taxon congruence in the rarity of subtidal rocky marine assemblages : no taxonomic shortcut for conservation monitoring. *Ecological Indicators* 77:239-249 [10.1016/j.ecolind.2017.02.012](https://doi.org/10.1016/j.ecolind.2017.02.012) ---  [HAL Link](https://HAL Link.archives-ouvertes.fr/HAL Link-01493202)

**Leroy B**, Meynard CN, Bellard C, Courchamp F. 2016. virtualspecies: an R package to generate virtual species distributions. *Ecography* 39:599-607 [10.1111/ecog.01388](https://doi.org/10.1111/ecog.01388)

**Leroy B**, Canard A & Ysnel F. 2013. Integrating multiple scales in rarity assessments of invertebrate taxa. Diversity and Distributions 19:794-803 [10.1111/ddi.12040](https://doi.org/10.1111/ddi.12040)

### Software

Lenormand M, **Leroy B**, Denelle P. 2023. bioregion: Comparison of Bioregionalisation Methods. Current version: 1.0  https://cran.r-project.org/package=bioregion Open-source code repository available here: https://github.com/bioRgeo/bioregion

**Leroy B.**, Kramer A, Vaissière AC, Diagne C. 2020-2023. **invacost**, package for the R software. Analyse Biological Invasion Costs with the 'InvaCost' Database. Published on the Comprehensive R Archive Network.  Current version: 1.1-1. https://cran.r-project.org/package=invacost Open-source code repository available here: https://github.com/Farewe/invacost 

**Leroy B.** 2019-2023. **biogeonetworks**, package for the R software. Biogeographical networks in R. Current version : 0.1.2. https://github.com/Farewe/biogeonetworks

**Leroy B.** 2014-2023. **virtualspecies**, package for the R software. A package to generate virtual species distribution for improvements of modeling techniques. Published on the Comprehensive R Archive Network. Current version : 1.5.1. http://borisleroy.com/virtualspecies

**Leroy B.** 2012-2023. **Rarity**, package for the R software. A package to calculate rarity indices for species and assemblages of species. Published on the Comprehensive R Archive Network. Current version : 1.3-6.  https://cran.r-project.org/package=Rarity


## Expertise
*I work at the science-policy interface, internationally (e.g., **Lead Author** for the European and Central Asia report by the IPBES) or specifically in France (e.g., conservation planning recommendations for the Terres Australes & Antarctiques Françaises).*

12 Expertise reports in total. Selected reports:
Ramiro-Sánchez B, **Leroy B**. 2022. *Progress report - Significant Adverse Impacts* Southern Indian Ocean Fisheries Agreement consultancy report. 38 pp.

Ramiro-Sánchez B, Martin A, **Leroy B**. 2022. *Final report on Consultancy on VME mapping.* Southern Indian Ocean Fisheries Agreement consultancy report. 33 pp.

Manfrini E, **Leroy B**, Diagne C, Soubeyran Y, Sarat E & Courchamp F. 2021. Les coûts économiques des invasions biologiques en France. Synthèse à l’intention des décideurs. Rapport MNHN/Université Paris Saclay/ IUCN, Paris, France.

Araignous E., Beaugrand G., Ben Rais Lasram F., Bourdaud P., Champagnat J.,  Dauvin J.-C., Grangeré K., HAL Linkouani G., Haraldsson M., Hattab T., Le Loc’h F., **Leroy B.**,  Lejart M., Niquil N., Nogues Q., Pagot J.-P., Raoux A., Safi, G. et Villanueva C. *Recommandations pour une approche écosystémique des aires d’implantation d’énergies marines renouvelables. Cas d’étude du parc éolien offshore de Courseulles-sur-mer. Plouzané*. France Energies Marines Editions, 2019, 60 pages. Edition : Juin 2019

**IPBES** (2018), *The IPBES regional assessment report on biodiversity and ecosystem services for Europe and Central Asia.* Rounsevell, M., Fischer, M., Torre-Marin Rando, A. and Mader, A. (eds.). Secretariat of the Intergovernmental Science-Policy Platform on Biodiversity and Ecosystem Services, Bonn, Germany. 892 pages. **Role: Lead Author (chapter 4), Contributing author (Chapters 1, 2, 3)**

## Research funding

`2021-2023` **1.5-year research project**, *Southern Indian Ocean
Fisheries Agreement*, Indian Ocean, 60 000€. PI. <br/> Bioregionalisation and 
Management of Vulnerable Marine Ecosystems in the Indian
Ocean.

`2021` **Expertise project**, *Planification de l'extension de la réserve Naturelle marine de Saint Paul et Amsterdam*, Terres Australes et Antarctiques Françaises, ~20k€.

`2020-2022`
**2-year research project**, *Southern Indian Ocean Fisheries Agreement*, Indian Ocean, 96 000€. PI.<br/> 
Mapping the biogeography of Vulnerable Marine Ecosystems in the Indian Ocean.

`2017–2019`
**2-year research project Labex**, *Muséum National d'Histoire Naturelle*, France, 100 000€. <br/>
Co-PI with Sarah Samadi & Maud Mouchet.<br/>
Valuing natural history collections and naturalist expeditions for the conservation and management of deep sea habitats: the Tropical Deep Sea Benthos program.

`2017–2019`
**2-year research project Labex**, *Muséum National d'Histoire Naturelle*, France, 100 000€.<br/> Co-PI with Sandrine Pavoine.<br/>
New indicators for biodiversity conservation based on phylogenetic originality and extinction risks : HEDGE and LEDGE species.

`2016-2019`
**3-year research project**, *ANR/France Energies Marines*, France. <br />
PI: Nathalie Niquil.<br/>
TROPHIK project: Ecosystemic approach of Renewable Marine Energies: Modelling the role of offshore windfarms in modifying coastal trophic network functionning and cumulated impacts of other drivers of change.


## Teaching

### Main courses

`2018-present`
*Master's degree, Muséum National d'Histoire Naturelle, Sorbonne Universités*, Paris, France.

- \- **Biogeography and Macroecology**, with Maud Mouchet, master's degree course, 10 days.

`2015-present`
*Doctoral school, Muséum National d'Histoire Naturelle, Sorbonne Universités*, Paris, France.

- \- **Species distribution models and climate change**, doctoral training school, 5 days.
- \- **Biogeography**, with Maud Mouchet, doctoral training school, 5 days.


### Short courses
Occasional short courses in the Muséum or elsewhere (e.g., Ecole Normale Supérieure)

### Online courses

`2020`
**The INVACOST R Package: Global Costs of Biological Invasions**
[Online tutorial link](https://github.com/Farewe/invacost)

`2019`
**Biogeographical networks in R**, [Online tutorial link](https://github.com/Farewe/biogeonetworks) 

`2014, 2019`
**The virtualspecies R package: a complete tutorial**, [Online tutorial link](http://borisleroy.com/virtualspecies_tutorial/)


## Supervision & evaluation

### Postdoctoral researchers

`2020-2022`
**Berta Ramiro-Sánchez**, *BOREA, Muséum National d'Histoire Naturelle*, Paris, France. <br />
Mapping the biogeography of Vulnerable Marine Ecosystems in the Indian Ocean.

`2018-2019`
**Lissette Victorero**, *BOREA, Muséum National d'Histoire Naturelle*, Paris, France. <br />
Co-supervised with Sarah Samadi & Maud Mouchet. <br />
Tropical Deep Sea Benthos project.

`2017-2019`
**Marine Robuchon**, *BOREA, Muséum National d'Histoire Naturelle*, Paris, France. <br />
Co-supervised with Sandrine Pavoine. <br />
Phylogenetic originality project.

`2017-2018`
**Juliette Delavenne**, *BOREA, Muséum National d'Histoire Naturelle*, Paris, France. <br /> Co-supervised with Sarah Samadi & Maud Mouchet. <br />
Tropical Deep Sea Benthos project.

`2017`
**Tarek Hattab**, *Oceanography & Geoscience lab, Université du Littoral Côté d'Opale*, Wimereux, France.
Co-supervised with Grégory Beaugrand. <br />
Trophik project.


### Doctoral students

`2022-2025`
**Eléna Manfrini**, *Ecology, Systematics, Evolution, Université Paris-Saclay*, Orsay, France.<br/>
Co-supervised with Franck Courchamp<br />
Risks of biological invasions from farmed arthropods.

`2019-2022`
**Camille Bernery**, *Ecology, Systematics, Evolution, Université Paris-Saclay*, Orsay, France.<br/>
Co-supervised with Franck Courchamp & Céline Bellard <br />
Biological invasions of freshwater fishes.

### Engineers

`2021`
**Andéol Bourgouin**, *Terres Australes et Antarctiques Françaises*, Paris, France.<br />
Bioregionalisation of the Saint-Paul Amsterdam EEZ in the Indian Ocean for the delineation of Marine Protected Areas.


### Master students

`2014-2022` 14 master students supervised in total (12 Master 2, 2 Master 1)

### Undergraduate interns

`2014-2022` 2 undergraduate internes supervised in total (1 Licence 3, 1 Licence 2)

### Committees and evaluations

`2017-2022` Examiner of 3 PhDs (2 in France, 1 in Australia)

`2014-2022` Member of 10 PhD committees (Comités de Suivi Individuel de thèse)

`2015-present`
**Evaluation committee**, *Master programs, Muséum National d'Histoire Naturelle, Sorbonne Universités*, Paris, France


## Professional service


`2010-present`
**Referee**. *Journals*: *African Journal of Ecology*, *Arachnologische Mitteilungen*, *Biological Invasions*, *Diversity & Distributions*, *Ecography*, *Ecological modelling*, *Ecology & Evolution*, *Ecology Letters*, *Environmental Modelling & Software*, *Freshwater Science*, *Frontiers in Ecology and the Environment*, *Global Ecology & Biogeography*, *Global Ecology & Conservation*, *International Journal of Geographic Information Science*, *Journal of Applied Ecology*, *Journal of Biogeography*, *Journal of Insect Conservation*, *PeerJ*, *PLoS one*, *Science China Life Sciences*, *Systematics & Biodiversity*.

**Research project proposals**: Agence Nationale pour la Recherche.

`2019-present`
**Associate Editor**. *Diversity & Distributions*

## Academic service

`2017-2019`
**Leader**, *group Biodiversity & Macroecology, Lab. Biology of Aquatic Organisms & Ecosystems, Muséum national d'Histoire Naturelle*, Paris, France.

## Outreach (examples)

`2021`
**Press** Multiple interviews on the themes of biological invasions, climate change and modelling (Science News, Polytechnique Insights, France TV info, Radio la 1ère//Planète Outre-Mer, Science & Vie)

`2017`
**Guest lectures**, *4th IMPRESS training workshop, Marie Sklodowska Curie Action Innovative Training network*, Chanteuges, France.
Two lectures given: **Aquatic biogeography and applications to conservation.** and **The Intergovernmental Science-Policy Platform on Biodiversity and Ecosystem Services (IPBES).**



```{r echo=FALSE, results="hide", message="hide", warning=FALSE} 
Sys.setlocale("LC_ALL","English")
```
<br/>Last updated: `r zoo::Sys.yearmon()`<br/><br/>
